### 2018-08-02 acknowledgements.tex
Backup and spell check

\section*{Acknowledgements}

The basis for this project stemmed from my eagerness to face new challenges in bioinformatics. For that reason, I am very grateful to Novozymes for having given me the opportunity to carry out my own project, and I really hope my results can be turned into innovative ways on how proteomics analyses are done in the company. Specially, I would like to thank Mathias Gruber for having supported and helped me throughout the journey and Anders M. Bernth Giessing for having come up with a project proposal on mass spectrometry and proteomics, a field of bioinformatics I knew little about at the project start. I would like to thank Samay Mir, Henrik M. Geertz-Hansen and Marija Majer for their suggestions and discussion.

Outside of Novozymes, I would like to thank my supervisor Thomas Hamelryck for his guidance and his feedback. Also, I would like to thank Harald Barsnes and Ludger Goeminne for their great advice and comments regarding the software implementation. Finally, I would like to mention those without whom the project would have been much harder to complete. My friends: Adham Khaled and Carlotta Porcelli and my family, thank you so much.