### 2018-11-22 figures/hierarch_diagram.tex
Clean repo

%% Hierarchical diagrams, DBDA
%% ---------------------------------
%% Tinu Schneider, October 2013
%
%
%\documentclass[12pt]{article}
%
%%\usepackage[T1]{fontenc} 
%%\usepackage[utf8]{inputenc}
%
%\usepackage[margin = 20mm, nohead, nofoot]{geometry} 
%
%\usepackage{tikz}
%    \usetikzlibrary{positioning, arrows, backgrounds, decorations, 
%    				decorations.pathmorphing, decorations.markings, calc}
%  
%\graphicspath{{./distribution_diagrams/svg_without_names}}
%    
%\pagestyle{empty}
%
%\usepackage{tgheros} 
%\renewcommand*\familydefault{\sfdefault} 
%	
%
%% Macros for the distributions
%% greek
%\newcommand{\normalGreek}[1]{\greek{#1}{normal} } 
%\newcommand{\gammaGreek}[1]{\greek{#1}{gamma} }
%\newcommand{\halfnormalGreek}[1]{\greek{#1}{halfnormal} }   
%
%
%\newcommand{\greek}[2]{\large$#1$\\ \includegraphics*[scale = 0.25]{distribution_diagrams/png_with_params/#2.png} }				 	 
%   			
%% letters (constants)  
%\newcommand{\normalLetters}[1]{\letters{#1}{normal}} 
%\newcommand{\gammaLetters} [1]{\letters{#1}{gamma}}   
%\newcommand{\halfnormalLetters} [1]{\letters{#1}{halfnormal}}   
%
%
%
%\newcommand{\letters}[2]{$#1$\\ \includegraphics*[scale = 0.25]{distribution_diagrams/png/#2.png} }
%
%\usepackage{stackengine, scalerel}
%\newcommand\Tunderbrace[3][]{%
%  \def\tmp{#2}%
%  \setbox0=\hbox{\tmp}%
%  \stackunder[1pt]{%
%    \stackunder[0pt]{\tmp}{\rotatebox{90}{\scaleto[2ex]{#1\{}{\wd0}}}%
%  }{%
%    \scriptsize #3%
%  }%
%}
%\newcommand\Toverbrace[3][]{%
%  \def\tmp{#2}%
%  \setbox0=\hbox{\tmp}%
%  \stackon[1pt]{%
%    \stackon[0pt]{\tmp}{\rotatebox{90}{\scaleto[2ex]{#1\}}{\wd0}}}%
%  }{%
%    \scriptsize #3%
%  }%
%}
%
%  						 	
%   						 	
%   						 	  						 	 
%\begin{document} % ------------------------------ %

% distances
\def\smallY{0.25}
\def\midY{2.25}
\def\bigY{4.5}

\def\smallX{1.2}
\def\midX{1.5}
\def\bigX{2.7}

\begin{center}
\begin{tikzpicture}[ framed, on grid,
   	node distance = \midY cm, 
   	inner frame xsep = 2.5ex,
   	inner frame ysep = 3.0ex,
    	box/.style = { rectangle,
	       		color = lightgray, 
	       		text = black,  
	       		text width = 2.0cm, 
	       		align = center},
	    boxFramed/.style = {box, 
	    		% draw, % uncomment this line, if you want the frames
	    		inner sep = 0pt},
	    boxFramedRound/.style = {boxFramed, rounded corners = 8pt},
	    boxFormula/.style = {box, inner sep = 1pt},
       	boxLegend/.style = {align = left, 
       			text width = 4cm, 
       			xshift = 2.2cm, 
       			yshift = 1.2cm,  
       			font = \scriptsize, 
       			color = black!70},
    	allArrows/.style = {font = \footnotesize, 
    			color = black!80, text = black},
    	arrSnake/.style = {allArrows, 
    			>=stealth', 
    			->, 
    			shorten >= 2pt, 
    			line join = round, 
    			decorate, 
  				decoration={snake, segment length = 10, 
  							amplitude = 1.2, post = lineto, 
  							post length = 4pt} },
     	implies/.style  = {allArrows, -implies, double, 
     			double equal sign distance, shorten >= 1pt}, 
     	labArrow/.style = {pos = 0.39},
     	labArrowLegend/.style = {labArrow, above, font = \tiny, inner sep = 3pt}
  	]

   	% nodes for the distributions and equation(s)
   	% starting at the bottom				
	\node [box, xshift = 2cm] (yi) {\Large $y_{i}$};

   	\node [boxFramedRound, above = of yi, yshift = -.2cm] (final) {\normalGreek{\mu_{i}, \epsilon} } ;

   	\node [boxFormula, above = of final, xshift=-\bigX cm, text width = 2.5 cm] (linMod)   
   			{ \large $\beta^0 +$\Toverbrace{$\beta^p_i + \beta^t_i + \beta^r_i$}{}} ;

   			   	\node [boxFormula, above = of linMod, xshift= 0 cm, text width = 2.5 cm] (selectPar)   
   			{ \large \Tunderbrace{$\beta^p \cdot x^p_i + \beta^t \cdot x^t_i + \beta^r \cdot x^r_i$}{} } ;

   	\node [boxFramed, above = of final,  xshift =  \bigX cm, yshift = \smallY cm] (proEpsilon) {\halfnormalGreek{S} } ;

   	\node [boxFramedRound, above = of selectPar, yshift=.2cm, xshift = -\bigX-2.4 cm]   (proIntercept)   {\normalGreek{M_0, S_0} };
   	\node [boxFramedRound, above = of selectPar, yshift=\smallY cm, xshift = \smallX cm]   (proBeta)    {\normalGreek{\mu_j, \sigma_j} } ;

   	\node [boxFramed, above = of proBeta,  xshift = -\smallX-1.2 cm] (promu)  {\normalGreek{M_j, \sigma_j} } ;
   	\node [boxFramed, above = of proBeta,  xshift =  \smallX+1.2 cm] (prosigma) {\halfnormalGreek{S_j} } ;

   	% Arrows	    	
   	\draw[arrSnake] (final.south) --  node[right, labArrow] {$\ i$} (yi);

    \draw[implies]  ($(linMod.south)+(1,-.15)$) -- node[right, labArrow] {$\ \ \ i$} (final.125);
    \draw[implies]  ($(selectPar.south)+(1.25,0)$) -- node[right, labArrow] {$\ i|p,t,r$} ($(linMod.100)+(1.1,0)$);

    \draw[arrSnake] (proEpsilon.south) -- (final.55);

    \draw[arrSnake] (proIntercept.south) -- node[left, labArrow] {$\ \ i$} ($(linMod)+(-1,0)$);       	 	
    \draw[arrSnake] (proBeta.south) --  node[left,  labArrow] {$j=p\quad $} ($(selectPar)+(-1,.8)$);
    \draw[arrSnake] (proBeta.south) --  node[xshift=0cm,  labArrow] {$j=t\quad $} ($(selectPar)+(1,.8)$); 
    \draw[arrSnake] (proBeta.south) --  node[right,  labArrow] {$j=r\quad $} ($(selectPar)+(2.5,.8)$);

    \draw[arrSnake] (promu.south) -- (proBeta.122);       	 	
    \draw[arrSnake] (prosigma.south) -- (proBeta.57);  
    \draw[arrSnake] (prosigma.west) -- (promu.east);

  	% Legend; needs improvement...
   	\node [boxLegend] at ( current bounding box.south west) (legendTitle) 
   		{%
   		\begin{tabular}{ll}
   			$A, \dots, Z$  & Constants \\
   			$a, \dots, z$ & Variables \\
   			$\alpha, \dots, \epsilon$ & Parameters \\ %[1ex]

			 & deterministic \\[-0.4ex]
			 & relation \\ %[1ex]

			 & stochastic \\[-0.4ex]
			 & relation \\ %[1ex]

			\multicolumn{2}{l}{$i = 1, \dots, N$} \\
			\multicolumn{2}{l}{$j \in$ peptide, treatment, run} \\

   		\end{tabular}
   	};

   	\def\xLegendArrowLeft{-3.75}
   	\def\xLegendArrowRight{-2.65}

   	\def\yLegendDouble{0.83}
   	\def\yLegendSnake{0.25}

   	\draw[implies, black!70]  (\xLegendArrowLeft, \yLegendDouble) -- (\xLegendArrowRight, \yLegendDouble);    	   	
    \draw[arrSnake, black!70] (\xLegendArrowLeft, \yLegendSnake) -- (\xLegendArrowRight, \yLegendSnake);

\end{tikzpicture}
\end{center}
%\end{document} % ------------------------------ %